Speeding up prediction performance of the BDT-based models

### 2017-11-11 acat_paper/paper.tex
Fixed TODO, removed paragraph about multicriteria optimisation

In this paper, we consider a modification of the training process of a predictive models by decision tree-based algorithm CatBoost [1] (TODO: 13).
\section{CatBoost}

@@

Let's consider another important training characteristic, namely the learning rate. The smaller it is, the smaller the contribution of each next tree in the ensemble. However, this has to be compensated by a large number of trees. The learning rate can be chosen by the cross-validation process given the fixed number of trees. Usually, the more trees the higher the model quality. (TODO: add conclusion)

@@

Let $B_i$ be the a of basket boundaries for the $i$-th feature, which was used in $T$. Then we can split the hyper-cube $[0;1]$ into small hyper-parallelepipeds. The prediction of the model is constant within each hyperparallelepiped. In fact, the model $Z$ allows you to calculate the value at any point of our hypercube. if we pre-calculate the value for each piece, then we will accelerate the application of our model, effectively reducing it to the time required to search a piece that contains the point. We called such approach a Decision Tensor. The asymptotic complexity of the application of the tree-based mode it $\mathcal(h\cdot t)$ where $h$ is the height of the trees, $t$ is the total number of the trees. Asymptotic complexity of Tensor is $\mathcal(n\cdot \log(b))$ where $n$ is the number of the features and $b$ is the number of cuts per feature (this complexity is estimated by using binary search). It should be noted for small $b$ simple linear search could be a bit faster that binary. So, with a small number of features the Tensor-based prediction can be computed much faster than usual tree-based model (TODO: 37)

@@

The consumption of memory for a single hypercube can be extremely huge provided that acceptable quality is achieved, so the natural extension of the idea is the construction of Decision Tensors ensemble. (TODO: 45)

@@

\section{Multicriteria optimization}
(TODO - probably should be removed)
The target metrics are $ Q $ (model quality), $ S $ (size), and $ t $ (time of inference). However, we can assume that the inference time is proportional to the number of decision tensors $ M $.
The input feature vectors will define the algorithm for obtaining the trees and the algorithm for combining them into the Tensors. We denote single vector by $ v $ and let $V$ be the set of all input vectors. Then we need to solve the following problem:

$$min(Q(V), S(V), M(V)), v \in V$$

It should be noted that for practical application this problem can be solved using the constrained optimization method. But in practice there are quite strict limitations on the size of the model and the available application time (TODO - what did you mean?).

>>>>

In this paper, we consider a modification of the training process of a predictive models by decision tree-based algorithm CatBoost [1]. But our approach is applicable to other boosting decision tree-based algorithms.

\section{CatBoost}

@@

Let's consider another important training characteristic, namely the learning rate. The smaller it is, the smaller the contribution of each next tree in the ensemble. However, this has to be compensated by a large number of trees and vice versa. The learning rate can be chosen by the cross-validation process given the fixed number of trees. Usually, the more trees the higher the model quality.

With default parameters CatBoost shows already good enough quality, so we variate only number of trees (and learning rate to compensate) because number of trees affects time. For other parameters was choosen default values.

@@

Let $B_i$ be the a of basket boundaries for the $i$-th feature, which was used in $T$. Then we can split the hyper-cube $[0;1]$ into small hyper-parallelepipeds. The prediction of the model is constant within each hyperparallelepiped. In fact, the model $Z$ allows you to calculate the value at any point of our hypercube. if we pre-calculate the value for each piece, then we will accelerate the application of our model, effectively reducing it to the time required to search a piece that contains the point. We called such approach a Decision Tensor. The asymptotic complexity of the application of the tree-based mode it $\mathcal(h\cdot t)$ where $h$ is the height of the trees, $t$ is the total number of the trees. Asymptotic complexity of Tensor is $\mathcal(n\cdot \log(b))$ where $n$ is the number of the features and $b$ is the number of cuts per feature (this complexity is estimated by using binary search). It should be noted for small $b$ simple linear search could be a bit faster that binary. So, with a small number of features the Tensor-based prediction may be computed much faster than tree-based model.

@@

The consumption of memory for a single tensor can be extremely huge provided that acceptable quality is achieved, so the natural extension of the idea is the construction of Decision Tensors ensemble but with some deterioration of applying time.

@@

% \section{Multicriteria optimization}
% (TODO - probably should be removed)
% The target metrics are $ Q $ (model quality), $ S $ (size), and $ t $ (time of inference). However, we can assume that the inference time is proportional to the number of decision tensors $ M $.
% The input feature vectors will define the algorithm for obtaining the trees and the algorithm for combining them into the Tensors. We denote single vector by $ v $ and let $V$ be the set of all input vectors. Then we need to solve the following problem:

% $$min(Q(V), S(V), M(V)), v \in V$$

% It should be noted that for practical application this problem can be solved using the constrained optimization method. But in practice there are quite strict limitations on the size of the model and the available application time (TODO - what did you mean?).
